---
title: "Permission Denied Stats in Freshwater ProbMon"
author: "Emma Jones"
date: "4/20/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(sf)
library(leaflet)
library(inlmisc)
library(readxl)
library(config)
library(pins)
library(gridExtra)

# Server connection things
conn <- config::get("connectionSettings") # get configuration settings


board_register_rsconnect(key = conn$CONNECT_API_KEY,  #Sys.getenv("CONNECT_API_KEY"),
                         server = conn$CONNECT_SERVER)#Sys.getenv("CONNECT_SERVER"))

wqmStations <- pin_get('ejones/WQM-Stations-Spatial', board = 'rsconnect') %>% 
  dplyr::select(StationID, Region = ASSESS_REG)

vahu6 <- st_read('C:/HardDriveBackup/R/GitHub/pinData/data/GIS/VA_SUBWATERSHED_6TH_ORDER_STG.shp') # this version of vahu6 layer goes outside state boundary
subbasinConversion <- read_csv('C:/HardDriveBackup/R/GitHub/pinData/data/subbasinToVAHU6conversion.csv')

subbasinLayer <- st_read('C:/HardDriveBackup/R/GitHub/IR2024/GIS/DEQ_VAHUSB_subbasins_EVJ.shp')  %>% 
  st_drop_geometry() %>% 
  rename('SUBBASIN' = 'SUBBASIN_1') %>% 
  distinct(subbasin, .keep_all = T) %>%  # don't need this by Assessment Region right now so simplify dataset to make next join cleaner
  dplyr::select(BASIN_NAME, BASIN_CODE, SUBBASIN, subbasin)

unrestrictedAssessmentRegionVAHU6Subbasin <- left_join(vahu6, subbasinConversion, by = c('VAHU6', 'VAHU5')) %>% 
  left_join(subbasinLayer, by = c('BASIN_CODE' = 'BASIN_CODE',
                                  'VAHUSB' = 'subbasin'))
#View(unrestrictedAssessmentRegionVAHU6Subbasin %>% st_drop_geometry())
```

This report seeks to answer a few questions about site permissions statewide using the Freshwater ProbMon dataset. This dataset is a great way to understand what is going on across the landscape, so though permission statistics aren't a metric one would typically use ProbMon data for, we can draw insights into permissions statewide, regionally, and temporally that might translate to other water monitoring programs.


The design status database for Freshwater ProbMon contains information on sites drawn vs sites sampled. We can use this to pull out different statistics on permission.

```{r}
original <- read.csv('C:/HardDriveBackup/R/GitHub/FreshwaterProbMonIntegratedReports/2022ProbChapter/processedData/designStatusIR2022.csv') %>% 
  mutate(StationID = sub("_.*$",'', sampleID)) %>% 
  left_join(wqmStations %>% distinct(StationID, .keep_all = TRUE), by = "StationID")

recent <- read_excel('C:/Users/wmu43954/OneDrive - Commonwealth of Virginia/Freshwater ProbMon/IR2024/Wadeable_ProbMon2021-2022.xlsx') %>% 
  dplyr::select(StationID, siteID, StationID_Trend, Region, `Latitude.DD` = Latitude, `Longitude.DD` = Longitude, status, Year) %>% 
  mutate(StationID = coalesce(StationID, siteID),
         sampleID = coalesce(StationID_Trend, StationID))

designStatus <- bind_rows(original, recent) %>%
  mutate(comment = as.character(comment)) %>% 
  st_as_sf(coords = c("Longitude.DD", "Latitude.DD"),  # make spatial layer using these columns
            remove = F, # don't remove these lat/lon cols from df
            crs = 4326)

# join region by spatial information
sfdesignStatus <- designStatus %>% 
  st_intersection(unrestrictedAssessmentRegionVAHU6Subbasin)

designStatus <- sfdesignStatus %>% st_drop_geometry()
```


First find out how many sites sampled. Have to be creative here bc boatable is coded differently

```{r}
boatable <- filter(designStatus, !is.na(status)) %>%
  filter(status %in% c('NT', "OT")) %>%
  filter(str_detect(comment, c("boatable", "Boatable", "TS-boatable",'boat', 'Bo') ) |
           str_detect(comment, c('Large', 'large')) |
           str_detect(comment, c('Boat site')) |
           str_detect(comment, "Boatable") |
           str_detect(comment, 'Large River') |
           str_detect(comment, 'Fall only - boat'))
wadeable <- filter(designStatus, status =='TS')

sampledSites <- rbind(wadeable, boatable)
```

How many sites have we sampled vs drawn in 2001-2022?

Sampled = `r nrow(sampledSites)` vs drawn `r nrow(designStatus)` 

What percent of sites have we sampled vs drawn in 2001-2022?

```{r}
nrow(sampledSites) / nrow(designStatus) * 100
```
72%


How many permission denied sites 2001-2022?

```{r}
nrow(filter(designStatus, status == 'PD'))
```

What percent permission denied compared to all sites 2001-2022?

```{r}
nrow(filter(designStatus, status == 'PD')) / nrow(filter(designStatus, !is.na(status))) * 100
```
roughly 7.5% across 2001-2022

How does that compare to other categories that were not sampled?

NT (dry, boatable, beavers, etc.)

```{r}
nrow(filter(designStatus, status == 'NT')) / nrow(filter(designStatus, !is.na(status))) * 100
```
29%

Our records on non sampled sites isn't great pre 2019, what was the % denial 2001-2018 and then 2019-2022?

```{r}
# 2001-2018
pre <- filter(designStatus, is.na(Year)) %>% 
  bind_rows(filter(designStatus, Year %in% 2001:2018))
nrow(filter(pre, status == 'PD')) / nrow(pre) * 100
preAnnualEstimate <-  nrow(filter(pre, status == 'PD')) / 18  

# # 2019
# nrow(filter(designStatus, Year == 2019 & status == 'PD')) / nrow(filter(designStatus, Year == 2019)) * 100
# # 2020
# nrow(filter(designStatus, Year == 2020 & status == 'PD')) / nrow(filter(designStatus, Year == 2020)) * 100
# # 2021
# nrow(filter(designStatus, Year == 2021 & status == 'PD')) / nrow(filter(designStatus, Year == 2021)) * 100
# # 2022
# nrow(filter(designStatus, Year == 2022 & status == 'PD')) / nrow(filter(designStatus, Year == 2022)) * 100

stats <- tibble( Year = 2001:2022) %>%  
                 # `Number Permission Denial` = as.numeric(NA), 
                 # `Pct Permission Denial` = NA) %>% 
  mutate(`Number Permission Denial` = case_when(Year %in% 2001:2018 ~ preAnnualEstimate,
                                                Year == 2019 ~ nrow(filter(designStatus, Year == 2019 & status == 'PD')) %>% as.numeric(),
                                                Year == 2020 ~ nrow(filter(designStatus, Year == 2020 & status == 'PD')) %>% as.numeric(),
                                                Year == 2021 ~ nrow(filter(designStatus, Year == 2021 & status == 'PD')) %>% as.numeric(),
                                                Year == 2022 ~ nrow(filter(designStatus, Year == 2022 & status == 'PD')) %>% as.numeric()),
         `Pct Permission Denial`  = case_when(Year %in% 2001:2018 ~ nrow(filter(pre, status == 'PD')) / nrow(pre) * 100,
                                                Year == 2019 ~ nrow(filter(designStatus, Year == 2019 & status == 'PD')) / nrow(filter(designStatus, Year == 2019)) * 100,
                                                Year == 2020 ~ nrow(filter(designStatus, Year == 2020 & status == 'PD')) / nrow(filter(designStatus, Year == 2020)) * 100,
                                                Year == 2021 ~ nrow(filter(designStatus, Year == 2021 & status == 'PD')) / nrow(filter(designStatus, Year == 2021)) * 100,
                                                Year == 2022 ~ nrow(filter(designStatus, Year == 2022 & status == 'PD')) / nrow(filter(designStatus, Year == 2022)) * 100),
         `BRRO Pct Permission Denial`  = case_when(Year %in% 2001:2018 ~ nrow(filter(pre, status == 'PD' & ASSESS_REG == 'BRRO')) / 
                                                     nrow(filter(pre, ASSESS_REG == 'BRRO')) * 100,
                                                   Year == 2019 ~ nrow(filter(designStatus, Year == 2019 & status == 'PD'  & ASSESS_REG == 'BRRO')) / 
                                                     nrow(filter(designStatus, Year == 2019 & ASSESS_REG == 'BRRO')) * 100,
                                                Year == 2020 ~ nrow(filter(designStatus, Year == 2020 & status == 'PD' & ASSESS_REG == 'BRRO')) /
                                                  nrow(filter(designStatus, Year == 2020 & ASSESS_REG == 'BRRO')) * 100,
                                                Year == 2021 ~ nrow(filter(designStatus, Year == 2021 & status == 'PD'  & ASSESS_REG == 'BRRO')) / 
                                                  nrow(filter(designStatus, Year == 2021 & ASSESS_REG == 'BRRO')) * 100,
                                                Year == 2022 ~ nrow(filter(designStatus, Year == 2022 & status == 'PD' & ASSESS_REG == 'BRRO')) /
                                                  nrow(filter(designStatus, Year == 2022 & ASSESS_REG == 'BRRO')) * 100),
         `PRO Pct Permission Denial`  = case_when(Year %in% 2001:2018 ~ nrow(filter(pre, status == 'PD' & ASSESS_REG == 'PRO')) / 
                                                     nrow(filter(pre, ASSESS_REG == 'PRO')) * 100,
                                                   Year == 2019 ~ nrow(filter(designStatus, Year == 2019 & status == 'PD'  & ASSESS_REG == 'PRO')) / 
                                                     nrow(filter(designStatus, Year == 2019 & ASSESS_REG == 'PRO')) * 100,
                                                Year == 2020 ~ nrow(filter(designStatus, Year == 2020 & status == 'PD' & ASSESS_REG == 'PRO')) /
                                                  nrow(filter(designStatus, Year == 2020 & ASSESS_REG == 'PRO')) * 100,
                                                Year == 2021 ~ nrow(filter(designStatus, Year == 2021 & status == 'PD'  & ASSESS_REG == 'PRO')) / 
                                                  nrow(filter(designStatus, Year == 2021 & ASSESS_REG == 'PRO')) * 100,
                                                Year == 2022 ~ nrow(filter(designStatus, Year == 2022 & status == 'PD' & ASSESS_REG == 'PRO')) /
                                                  nrow(filter(designStatus, Year == 2022 & ASSESS_REG == 'PRO')) * 100),
         `NRO Pct Permission Denial`  = case_when(Year %in% 2001:2018 ~ nrow(filter(pre, status == 'PD' & ASSESS_REG == 'NRO')) / 
                                                     nrow(filter(pre, ASSESS_REG == 'NRO')) * 100,
                                                   Year == 2019 ~ nrow(filter(designStatus, Year == 2019 & status == 'PD'  & ASSESS_REG == 'NRO')) / 
                                                     nrow(filter(designStatus, Year == 2019 & ASSESS_REG == 'NRO')) * 100,
                                                Year == 2020 ~ nrow(filter(designStatus, Year == 2020 & status == 'PD' & ASSESS_REG == 'NRO')) /
                                                  nrow(filter(designStatus, Year == 2020 & ASSESS_REG == 'NRO')) * 100,
                                                Year == 2021 ~ nrow(filter(designStatus, Year == 2021 & status == 'PD'  & ASSESS_REG == 'NRO')) / 
                                                  nrow(filter(designStatus, Year == 2021 & ASSESS_REG == 'NRO')) * 100,
                                                Year == 2022 ~ nrow(filter(designStatus, Year == 2022 & status == 'PD' & ASSESS_REG == 'NRO')) /
                                                  nrow(filter(designStatus, Year == 2022 & ASSESS_REG == 'NRO')) * 100),
         `VRO Pct Permission Denial`  = case_when(Year %in% 2001:2018 ~ nrow(filter(pre, status == 'PD' & ASSESS_REG == 'VRO')) / 
                                                     nrow(filter(pre, ASSESS_REG == 'VRO')) * 100,
                                                   Year == 2019 ~ nrow(filter(designStatus, Year == 2019 & status == 'PD'  & ASSESS_REG == 'VRO')) / 
                                                     nrow(filter(designStatus, Year == 2019 & ASSESS_REG == 'VRO')) * 100,
                                                Year == 2020 ~ nrow(filter(designStatus, Year == 2020 & status == 'PD' & ASSESS_REG == 'VRO')) /
                                                  nrow(filter(designStatus, Year == 2020 & ASSESS_REG == 'VRO')) * 100,
                                                Year == 2021 ~ nrow(filter(designStatus, Year == 2021 & status == 'PD'  & ASSESS_REG == 'VRO')) / 
                                                  nrow(filter(designStatus, Year == 2021 & ASSESS_REG == 'VRO')) * 100,
                                                Year == 2022 ~ nrow(filter(designStatus, Year == 2022 & status == 'PD' & ASSESS_REG == 'VRO')) /
                                                  nrow(filter(designStatus, Year == 2022 & ASSESS_REG == 'VRO')) * 100),
         `SWRO Pct Permission Denial`  = case_when(Year %in% 2001:2018 ~ nrow(filter(pre, status == 'PD' & ASSESS_REG == 'SWRO')) / 
                                                     nrow(filter(pre, ASSESS_REG == 'SWRO')) * 100,
                                                   Year == 2019 ~ nrow(filter(designStatus, Year == 2019 & status == 'PD'  & ASSESS_REG == 'SWRO')) / 
                                                     nrow(filter(designStatus, Year == 2019 & ASSESS_REG == 'SWRO')) * 100,
                                                Year == 2020 ~ nrow(filter(designStatus, Year == 2020 & status == 'PD' & ASSESS_REG == 'SWRO')) /
                                                  nrow(filter(designStatus, Year == 2020 & ASSESS_REG == 'SWRO')) * 100,
                                                Year == 2021 ~ nrow(filter(designStatus, Year == 2021 & status == 'PD'  & ASSESS_REG == 'SWRO')) / 
                                                  nrow(filter(designStatus, Year == 2021 & ASSESS_REG == 'SWRO')) * 100,
                                                Year == 2022 ~ nrow(filter(designStatus, Year == 2022 & status == 'PD' & ASSESS_REG == 'SWRO')) /
                                                  nrow(filter(designStatus, Year == 2022 & ASSESS_REG == 'SWRO')) * 100),
         `TRO Pct Permission Denial`  = case_when(Year %in% 2001:2018 ~ nrow(filter(pre, status == 'PD' & ASSESS_REG == 'TRO')) / 
                                                     nrow(filter(pre, ASSESS_REG == 'TRO')) * 100,
                                                   Year == 2019 ~ nrow(filter(designStatus, Year == 2019 & status == 'PD'  & ASSESS_REG == 'TRO')) / 
                                                     nrow(filter(designStatus, Year == 2019 & ASSESS_REG == 'TRO')) * 100,
                                                Year == 2020 ~ nrow(filter(designStatus, Year == 2020 & status == 'PD' & ASSESS_REG == 'TRO')) /
                                                  nrow(filter(designStatus, Year == 2020 & ASSESS_REG == 'TRO')) * 100,
                                                Year == 2021 ~ nrow(filter(designStatus, Year == 2021 & status == 'PD'  & ASSESS_REG == 'TRO')) / 
                                                  nrow(filter(designStatus, Year == 2021 & ASSESS_REG == 'TRO')) * 100,
                                                Year == 2022 ~ nrow(filter(designStatus, Year == 2022 & status == 'PD' & ASSESS_REG == 'TRO')) /
                                                  nrow(filter(designStatus, Year == 2022 & ASSESS_REG == 'TRO')) * 100)
         )
         
         
```

Statewide plots

```{r}
ggplot(stats, aes(x = Year, y = `Pct Permission Denial`))+
  geom_bar(stat="identity")+
  ggtitle("Statewide Percent Permission Denied By Year")+
  theme_bw()
```


```{r}
ggplot(stats, aes(x = Year, y = `Number Permission Denial`))+
  geom_bar(stat="identity")+
  ggtitle("Statewide Number of Sites Permission Denied By Year")+
  theme_bw()
```


Percent Permission Denied by Region

```{r}
BRRO <- ggplot(stats, aes(x = Year, y = `BRRO Pct Permission Denial`))+
  geom_bar(stat="identity")+
  ylab("BRRO % PD")+
  #ggtitle("BRRO Percent of Sites Permission Denied By Year")+
  theme_bw()
PRO <- ggplot(stats, aes(x = Year, y = `PRO Pct Permission Denial`))+
  geom_bar(stat="identity")+
    ylab("PRO % PD")+

  #ggtitle("PRO Percent of Sites Permission Denied By Year")+
  theme_bw()
NRO <- ggplot(stats, aes(x = Year, y = `NRO Pct Permission Denial`))+
  geom_bar(stat="identity")+
    ylab("NRO % PD")+

  #ggtitle("NRO Percent of Sites Permission Denied By Year")+
  theme_bw()
VRO <- ggplot(stats, aes(x = Year, y = `VRO Pct Permission Denial`))+
  geom_bar(stat="identity")+
    ylab("VRO % PD")+

  #ggtitle("VRO Percent of Sites Permission Denied By Year")+
  theme_bw()
SWRO <- ggplot(stats, aes(x = Year, y = `SWRO Pct Permission Denial`))+
  geom_bar(stat="identity")+
    ylab("SWRO % PD")+

  #ggtitle("VRO Percent of Sites Permission Denied By Year")+
  theme_bw()
TRO <- ggplot(stats, aes(x = Year, y = `TRO Pct Permission Denial`))+
  geom_bar(stat="identity")+
    ylab("TRO % PD")+

  #ggtitle("VRO Percent of Sites Permission Denied By Year")+
  theme_bw()


grid.arrange(arrangeGrob(BRRO, 
                         NRO,
                         PRO,
                         VRO,
                         SWRO, TRO,
                         nrow = 3),    nrow=1)
```


Statewide map of sites that were permission denied 2001-2022:

```{r}
library(leaflet)
library(inlmisc)

assessmentRegions <- st_as_sf(pin_get('ejones/AssessmentRegions_simple', board = 'rsconnect'))
basins <- unrestrictedAssessmentRegionVAHU6Subbasin %>% 
  group_by(BASIN_NAME) %>% summarise()

 # color palette for assessment polygons
  pal <- colorFactor(
    palette = topo.colors(7),
    domain = assessmentRegions$ASSESS_REG)
PD <- filter(sfdesignStatus, status == 'PD') 

CreateWebMap(maps = c("Topo","Imagery","Hydrography"), collapsed = TRUE, 
                 options= leafletOptions(zoomControl = TRUE,minZoom = 3, maxZoom = 20,
                                         preferCanvas = TRUE)) %>%
      setView(-78, 37.5, zoom=7)  %>% 
      addPolygons(data= assessmentRegions,  color = 'black', weight = 1,
                  fillColor= ~pal(assessmentRegions$ASSESS_REG), fillOpacity = 0.5,stroke=0.1,
                  group="Assessment Regions",
                  label = ~ASSESS_REG, layerId = ~ASSESS_REG)  %>% hideGroup('Assessment Regions') %>% 
  addPolygons(data= basins,  color = 'black', weight = 1,
                  fillColor= ~pal(basins$BASIN_NAME), fillOpacity = 0.5,stroke=0.1,
                  group="Major Basins",
                  label = ~BASIN_NAME, layerId = ~BASIN_NAME) %>% hideGroup('Major Basins') %>% 
  addCircleMarkers(data = PD, color='blue', fillColor='yellow', radius = 4,
                       fillOpacity = 0.5,opacity=0.8,weight = 2,stroke=T, group="Permission Denied Sites",
                       label = ~sampleID, layerId = ~sampleID) %>%
      addLayersControl(baseGroups=c("Topo","Imagery","Hydrography"),
                       overlayGroups = c('Permission Denied Sites',
                                         "Major Basins",
                                         'Assessment Regions'),
                       options=layersControlOptions(collapsed=T),
                       position='topleft')
```

